---
title: "Extract to create a csv file used to feed FGP/OGSL/ETC"
author: "Pablo Vergara"
date: "February 29 2024"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
This document is the narrative of how an Excel file called "sGSL-September-RV-FGP.csv" was generated.

The DFO Gulf Region September RV survey follows a stratified random sampling design and covers Division 4T of the Northwest Atlantic Fisheries Organisation (NAFO).


```{r setup, fig.height=6, fig.width=6}
`%nin%` = Negate(`%in%`) ## a useful operator
suppressPackageStartupMessages(library(gulf))
```

## SETS
Using the function "read.card" from the DFO Gulf Region's R package "gulf", read in the set cards and exclude hydrographic stations and null sets, so as to keep only representative tows in strata 401 to 403 and 415 to 439.

```{r sets}
yrs <- 1971:2023

projectBase = here::here()
# source('D:/SourceControl/Git Hub/gulf/R/rv.good.sets.R')
# x <- read.card(card.type="set", year=yrs)
x <-  rv.good.sets(year=yrs, source="oracle")
## we want to keep only representative tows, within the historical context that some tows were identified as "hydrography" before 1993 but were really fishing stations
uu <- sort(unique(x$experiment))
```

```{r results='asis'}
knitr::kable(data.frame(uu,experiment.str(uu)))## what are the experiment codes
```

```{r}
#Need to remove data that is of no use for analysis

# true for sets other than NULL sets
#  logical1 <- (x$experiment != 3)
# # 
# # # true for all except 2 tows in 1982 that were true hydrography set cards
#  logical2 <- !((x$year==1982 & x$cruise.number==278) & x$experiment %in% c(8,9))
# # 
# # ## true for all except true hydrography set cards after 1993
#  logical3 <- !(x$year>1993 & x$experiment %in% c(8,9))


## keep only the strata in NAFO 4T,
logical4 <- x$stratum %in% c(401,402,403,415:439)


#Subset set cards with conditions
x <- x[logical4,]
```

Add a few useful columns to the data frame containing the set card information (including depth and swept area as requested).
```{r sets2}
## add useful columns
x[x$vessel.code=="T" & x$year==2003,"vessel.code"] <- "TE" ## CCGS Templeman used in 2003
x$unique.id <- paste(x$year, x$cruise.number, x$vessel.code, x$set.number, sep="-")
x$experiment.str <- experiment.str(x$experiment)
x$vessel.str <- vessel.str(x$vessel.code)
x$gear.str <- gear.str(x$gear)
x$longitude <- longitude(x)
x$latitude <- latitude(x)
x$mission <- paste(x$vessel.code, x$cruise.number, sep="")

ox <- order(x$year, x$month, x$day, x$start.hour)
x <- x[ox,] # reorder chronologically

## remove missions N176 and H245 which were comparative missions conducted in August 1992
x <- x[which(x$mission !="N176"),]
x <- x[which(x$mission !="H245"),]

```

To deal with "repeat tows" (fishing locations that were purposefully sampled more than once in a given year) an additional column called station.number is added to identify tows that were conducted at the same location within a survey and that should be treated differently when estimating species density. These repeating tows are not independent observations and should be treated accordingly in analyses.

```{r sets3}
x$station.number <- unlist(sapply(yrs, function(y){x.t <- x[x$year==y,]; station.number(x.t, method="observed")}))
```

Catch cards contain the total catch information for the species of interest. Here they are adjusted for distance towed, estimated diurnal effects and estimated vessel-gear effects.
```{r catch2}
### Catch card for all years requested (1970 - 20252)
y <- read.card(card.type="catch", year = yrs)
```

## CATCH
```{r adjust}

y[y$vessel.code=="T" & y$year==2003,"vessel.code"] <- "TE" ## CCGS Templeman used in 2003

# y <- adjust(y, x)
y$unique.id <- paste(y$year, y$cruise.number, y$vessel.code, y$set.number, sep="-")

uni.spec = unique(y$species)
columns = c("species","english.name", "french.name", "latin.name")

 
 
#Create a Empty DataFrame with length(uni.spec) rows and length(columns) columns
df = data.frame(matrix(nrow = length(uni.spec), ncol = length(columns))) 
colnames(df) = columns
df$species = uni.spec
df$english.name <- species.str( uni.spec, "english")
df$latin.name   <- species.str( uni.spec, "latin")
df$french.name  <- species.str( uni.spec, "french")


y <- merge(y, df, by = "species", names = c("english.name","french.name", "latin.name"))
z <- merge(y,x, all.x = TRUE, by = "unique.id", names = c("longitude","latitude", "gear.str"))

##  CSV
## write catch cards to file
fn2 <- "sGSL-September-RV-FGP.csv"
fp  <-paste0(projectBase,"/stock_assessment_surveys/1989de32 (rv_survey)/")
       
csv.fn2 <- file.path(fp,fn2)

ooz <- order(z$year, z$month, z$day, z$start.hour, z$start.minute, z$species)

#columns to keep
fvars2 <- c("year","month","day","start.hour","start.minute","latitude","longitude","gear.str","species","french.name","english.name","latin.name","weight.caught","number.caught")
zz=  z[ooz,fvars2]

#header for FGP must be in both languages
names(zz) <- c("year__annee","month__mois","day__jour","start_hour__heure_de_depart","start_minute__minute_de_depart","latitude","longitude","gear__equipement","species__espece","french_name__nom_francais","english_name__nom_anglais","latin_name__nom_latin","weight_caught__poids_pris","number_caught__quantite_attrape")

#fixes encoding issue
con<-file(csv.fn2,encoding="ISO-8859-1")
write.csv(zz, file=con, row.names=FALSE)
```
