---
title: "Extract to create a csv file used to feed FGP/OGSL/ETC"
author: "Pablo Vergara"
date: "March 7 2023"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
This document is the narrative of how an Excel file called "sGSL-scallop-RV-FGP.csv" was generated.


```{r setup, fig.height=6, fig.width=6}
`%nin%` = Negate(`%in%`) ## a useful operator
suppressPackageStartupMessages(library(gulf))
```

## SETS
Using the function "read.card" from the DFO Gulf Region's R package "gulf", read in the set cards and exclude hydrographic stations and null sets, so as to keep only representative tows.

```{r sets}
survey = "sca"
projectBase = here::here()
 yrs <- c(2012, 2013, 2014, 2015, 2016,2018,2019,2020,2021, 2022)

x <- read.card(card.type="set", year=yrs, survey = survey)
## we want to keep only representative tows, within the historical context that some tows were identified as "hydrography" before 1993 but were really fishing stations
uu <- sort(unique(x$experiment))
```

```{r results='asis'}
knitr::kable(data.frame(uu,experiment.str(uu)))## what are the experiment codes
```

```{r}
#Need to remove data that is of no use for analysis

# true for sets other than NULL sets
logical1 <- x$experiment == 1

#Subset set cards with conditions
x <- x[which(logical1),]
```

Add a few useful columns to the data frame containing the set card information (including depth and swept area as requested).
```{r sets2}
## add useful columns
x$unique.id <- paste(x$year, x$cruise.number, x$vessel.code, x$set.number, sep="-")
x$experiment.str <- experiment.str(x$experiment)
x$vessel.str <- vessel.str(x$vessel.code)
x$gear.str <- gear.str(x$gear)
x$longitude <- longitude(x)
x$latitude <- latitude(x)
x$mission <- paste(x$vessel.code, x$cruise.number, sep="")

ox <- order(x$year, x$month, x$day, x$start.hour)
x <- x[ox,] # reorder chronologically

```



Catch cards contain the total catch information for the species of interest. Here they are adjusted for distance towed, estimated diurnal effects and estimated vessel-gear effects.
```{r catch2}
### Catch card for all years requesteed
y <- read.card(card.type="cat", year = yrs , survey=survey)

```

## CATCH
```{r adjust}
# y <- adjust(y, x)
tmp <- adjust(y, x)
y$unique.id <- paste(y$year, y$cruise.number, y$vessel.code, y$set.number, sep="-")
y$english.name <- species.str(y$species, "english")
y$latin.name <- species.str(y$species, "latin")
y$french.name <- species.str(y$species, "french")

z <- merge(y,x, all.x = TRUE, by = "unique.id", names = c("longitude","latitude", "gear.str"))

##  CSV
## write to file
fn2 <- "sGSL-scallop-RV-FGP.csv"
fp  <- paste0(projectBase,"/stock_assessment_surveys/de6dc782 (scallop_survey)/")
csv.fn2 <- file.path(fp,fn2)

ooz <- order(z$year, z$month, z$day, z$start.hour, z$start.minute, z$species)

#columns to keep
fvars2 <- c("year","month","day","start.hour","start.minute","latitude","longitude","gear.str","species","french.name","english.name","latin.name","weight.caught","number.caught")


zz=  z[ooz,fvars2]
#rename header for FGP
 
names(zz) <- c("year__annee","month__mois","day__jour","start_hour__heure_de_depart","start_minute__minute_de_depart","latitude","longitude","gear__equipement","species__espece","french_name__nom_francais","english_name__nom_anglais","latin_name__nom_latin","weight_caught__poids_pris","number_caught__quantite_attrape")

write.csv(zz, file=csv.fn2, row.names=FALSE)
```
