---
title: "Gulf Fisheries Centre Ageing Structures collection"
author: "Daniel Ricard"
date: '`r paste0("Last modified timestap: ", format(Sys.time(), "%Y-%m-%d %H:%M:%S"))`'
output: html_notebook
---

The Ageing Structures collection held at the Gulf Fisheries Centre consists in Teleost saggital otoliths and Chondrichthyan vertebrae. For certain species, the physical specimens held in the collection are often accompanied with digital images. The individual details associated with each structure consists in capture date and location, length of individual and usually includes additional variables such as weight and maturity status. 

The purpose of this notebook is to facilitate the generation and update of the FGP datasets associated with the Ageing Structure collection.

```{r loaddata}
library(gulf)
data(rv)
```

## Data dictionary
```{r dictionary}
#################################################################
#################################################################
## data dictionary including the number of aged otoliths
vars.bilingual <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")

data.dic <- data.frame(
  name_nom = vars.bilingual
  )
data.dic$description_fr <- c("Source des otolithes, peut soit provenir du relevé par navire de recherche dans la division 4T de l'OPANO (Research Vessel (RV-4T)), du relevé par navire de recherche dans les divisions 4RS de l'OPANO (Research Vessel (RV-4RS)), du relevé sentinelle de la division 4T de l'OPANO (SENT-4T),  d'échantillonnage des pêches commerciales (Commercial) ou d'observateurs en mer (Observateurs)", "Nom latin de l'espèce", "Nom anglais de l'espèce", "Nom français de l'espèce", "Année d'échantillonnage des otolithes", "Mois d'échantillonnage des otolithes", "Nombre d'otolithes dans la collection", "Nombre d'otolithes qui ont été agées")

data.dic$description_en <- c("Source of otoliths, can be either from the NAFO division 4T Research Vessel (RV-4T) survey, the NAFO divisions 4RS Research Vessel (RV-4RS) survey, the NAFO division 4T sentinel survey (SENT-4T), from commercial sampling (Commercial) or from at-sea observers (Observers)", "Latin name of the species", "English name of the species", "French name of the species", "Year when otoliths were collected", "Month when otolith were collected", "Number of otoliths in collection", "Number of otoliths that were aged")

data.dic.fn <- "data-dictionary.csv"
write.csv(data.dic, file=data.dic.fn, row.names = FALSE, fileEncoding = "UTF8")

#################################################################
## data dictionary for skates
vars.bilingual <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.vertebrae_nombre.vertèbres")

data.dic <- data.frame(
  name_nom = vars.bilingual
  )
data.dic$description_fr <- c("Source des vertèbres, le relevé par navire de recherche (Research Vessel (RV))", "Nom latin de l'espèce", "Nom anglais de l'espèce", "Nom français de l'espèce", "Année d'échantillonnage des vertèbres", "Mois d'échantillonnage des vertèbres", "Nombre de vertèbres dans la collection")

data.dic$description_en <- c("Source of vertebrae, the Research Vessel (RV) survey", "Latin name of the species", "English name of the species", "French name of the species", "Year when vertebrae were collected", "Month when vertebrae were collected", "Number of vertebrae in collection")

data.dic.fn <- "data-dictionary-skates.csv"
write.csv(data.dic, file=data.dic.fn, row.names = FALSE, fileEncoding = "UTF8")
```


## Atlantic Cod
Otoliths from the September RV survey, from other surveys, from sentinel programs, from port sampling of commercial fisheries and from at-sea observers.

WHERE ARE THE COMMERCIAL SAMPLES FOR 2020 AND 2021?

```{r cod, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species==10),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV-4T"

## Sentinel data
y.s <- x <- read.card("sentinel", species=10, card.type="bio")
sent.df <- aggregate(fish.number~year+month, data=y.s[which(y.s$age.material==1),], length)
names(sent.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
sent.df$source <- "SENT-4T"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="codage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=10, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in c(1983:2002,2004:2019,2021)){ ## loop over years 
t.c <- gulf::read.card(year=yy, sampling="commercial", species=10, card.type="age")
y.c <- rbind(y.c,t.c)
}
l.4tvn <- which(substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") ## keep only otoliths collected in NAFO 4TVN
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

## observers
y.o <- read.card(sampling = "obs", year = 1994:2020, card = "age", species = 10)
obs.df <- aggregate(fish.number~year+month, data=y.o, length)
names(obs.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
obs.df$source <- "Observers"

out.df <- rbind(rv.df, comm.df, obs.df, sent.df)
out.df$latin.name_nom.latin <- "Gadus morhua"
out.df$english.name_nom.anglais <- "Atlantic Cod"
out.df$french.name_nom.français <- "Morue franche"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4TVN-Atlantic-Cod-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")


## how many were actually aged?
## RV
y <- rv$bio[which(rv$bio$species==10),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.rv.df$source <- "RV-4T"
m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## Sentinel
sent.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(sent.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.sent.df <- merge(sent.df, sent.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.sent.df$source <- "SENT-4T"
m.sent.df[is.na(m.sent.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0


## do the same for commercial samples, how many were aged
l.4tvn <- which((substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") & y.c$age %in% c(0:25)) 
comm.df2 <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.comm.df <- merge(comm.df, comm.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.comm.df$source <- "Commercial"
m.comm.df[is.na(m.comm.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## observers
l.aged <- which(y.o$age %in% c(0:25)) 
obs.df2 <- aggregate(fish.number~year+month, data=y.o[l.aged,], length)
names(obs.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.obs.df <- merge(obs.df, obs.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.obs.df$source <- "Observers"
m.obs.df[is.na(m.obs.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

out.df <- rbind(m.rv.df, m.comm.df, m.obs.df, m.sent.df)
out.df$latin.name_nom.latin <- "Gadus morhua"
out.df$english.name_nom.anglais <- "Atlantic Cod"
out.df$french.name_nom.français <- "Morue franche"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4TVN-Atlantic-Cod-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

```


## White Hake
Otoliths from the September RV survey, from other surveys, from sentinel programs, from port sampling of commercial fisheries and from at-sea observers.

WHERE ARE THE COMMERCIAL SAMPLES FOR 2020 AND 2021?

```{r hake, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species==12),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV-4T"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="whage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=12, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in c(1983:2019)){ ## loop over years
  print(paste0("Year ",yy))
t.c <- gulf::read.card(year=yy, sampling="commercial", species=12, card.type="age")
y.c <- rbind(y.c,t.c)
}
l.4tvn <- which(substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") ## keep only otoliths collected in NAFO 4TVN
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

## observers
y.o <- read.card(sampling = "obs", year = 2000:2020, card = "age", species = 12)
obs.df <- aggregate(fish.number~year+month, data=y.o, length)
names(obs.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
obs.df$source <- "Observers"


out.df <- rbind(rv.df, comm.df,obs.df)
out.df$latin.name_nom.latin <- "Urophycis tenuis"
out.df$english.name_nom.anglais <- "White Hake"
out.df$french.name_nom.français <- "Merluche blanche"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-White-Hake-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

## how many were actually aged?
y <- rv$bio[which(rv$bio$species==12),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.rv.df$source <- "RV-4T"
m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## do the same for commercial samples, how many were aged
l.4tvn <- which((substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") & y.c$age %in% c(0:25)) 
comm.df2 <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.comm.df <- merge(comm.df, comm.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.comm.df$source <- "Commercial"
m.comm.df[is.na(m.comm.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## observers
l.aged <- which(y.o$age %in% c(0:25)) 
obs.df2 <- aggregate(fish.number~year+month, data=y.o[l.aged,], length)
names(obs.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.obs.df <- merge(obs.df, obs.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.obs.df$source <- "Observers"
m.obs.df[is.na(m.obs.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0


out.df <- rbind(m.rv.df, m.comm.df, m.obs.df)
out.df$latin.name_nom.latin <- "Urophycis tenuis"
out.df$english.name_nom.anglais <- "White Hake"
out.df$french.name_nom.français <- "Merluche blanche"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-White-Hake-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

```

## American Plaice
Otoliths from the September RV survey, from other surveys, from sentinel programs, from port sampling of commercial fisheries and from at-sea observers.

WHERE ARE THE COMMERCIAL SAMPLES FOR 2020 AND 2021?

```{r plaice, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species==40),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV-4T"
## oo <- order(rv.df$year, rv.df$month)
## rv.df[oo,]

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="plaage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=40, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in 1983:2019){ ## loop over years
  print(paste0("Year ",yy))
t.c <- gulf::read.card(year=yy, sampling="commercial", species=40, card.type="age")
y.c <- rbind(y.c,t.c)
}

l.4t <- which(substring(y.c$nafo.division,1,2)=="4T") ## keep only otoliths collected in NAFO 4T

comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

## observers data
y.o <- read.card(sampling = "obs", year = 2000:2020, card = "age", species = 40)
obs.df <- aggregate(fish.number~year+month, data=y.o, length)
names(obs.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
obs.df$source <- "Observers"


out.df <- rbind(rv.df, comm.df, obs.df)
out.df$latin.name_nom.latin <- "Hippoglossoides platessoides"
out.df$english.name_nom.anglais <- "American Plaice"
out.df$french.name_nom.français <- "Plie canadienne"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-American-Plaice-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

## how many were actually aged?
y <- rv$bio[which(rv$bio$species==40),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.rv.df$source <- "RV-4T"
m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## do the same for commercial samples, how many were aged
l.4tvn <- which((substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") & y.c$age %in% c(0:25)) 
comm.df2 <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.comm.df <- merge(comm.df, comm.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.comm.df$source <- "Commercial"
m.comm.df[is.na(m.comm.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## observers
l.aged <- which(y.o$age %in% c(0:25)) 
obs.df2 <- aggregate(fish.number~year+month, data=y.o[l.aged,], length)
names(obs.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.obs.df <- merge(obs.df, obs.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.obs.df$source <- "Observers"
m.obs.df[is.na(m.obs.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

out.df <- rbind(m.rv.df, m.comm.df, m.obs.df)
out.df$latin.name_nom.latin <- "Hippoglossoides platessoides"
out.df$english.name_nom.anglais <- "American Plaice"
out.df$french.name_nom.français <- "Plie canadienne"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-American-Plaice-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

```

## Witch Flounder
Otoliths from the September RV survey, from other surveys, from sentinel programs and from port sampling of commercial fisheries and from at-sea observers. Otoliths collected in the Quebec August survey in 4RS.

WHERE ARE THE COMMERCIAL SAMPLES FOR 2017 TO 2021?


```{r witch, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species ==41),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV-4T"

## oo <- order(rv.df$year, rv.df$month)
## rv.df[oo,]
## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="witage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=41, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in 1998:2019){ ## loop over years
  print(paste0("Year ",yy))
t.c <- gulf::read.card(year=yy, sampling="commercial", species=41, card.type="age")
y.c <- rbind(y.c,t.c)
}

l.4t <- which(substring(y.c$nafo.division,1,2) %in% c("4R","4S","4T")) ## keep only otoliths collected in NAFO 4RST
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

## observers data
y.o <- read.card(sampling = "obs", year = 1999:2020, card = "age", species = 41)
obs.df <- aggregate(fish.number~year+month, data=y.o, length)
names(obs.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
obs.df$source <- "Observers"


out.df <- rbind(rv.df, comm.df, obs.df)
out.df$latin.name_nom.latin <- "Glyptocephalus cynoglossus"
out.df$english.name_nom.anglais <- "Witch Flounder"
out.df$french.name_nom.français <- "Plie grise"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4RST-Witch-Flounder-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

## how many were actually aged?
y <- rv$bio[which(rv$bio$species==41),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.rv.df$source <- "RV-4T"
m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## do the same for commercial samples, how many were aged
l.4tvn <- which((substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") & y.c$age %in% c(0:25)) 
comm.df2 <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.comm.df <- merge(comm.df, comm.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.comm.df$source <- "Commercial"
m.comm.df[is.na(m.comm.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## do the same for observer samples, how many were aged
l.4tvn <- which(y.o$age %in% c(0:25)) 
m.obs.df <- obs.df
m.obs.df$number.otoliths.aged_nombre.otolithes.agées <- 0
  
out.df <- rbind(m.rv.df, m.comm.df, m.obs.df)
out.df$latin.name_nom.latin <- "Glyptocephalus cynoglossus"
out.df$english.name_nom.anglais <- "Witch Flounder"
out.df$french.name_nom.français <- "Plie grise"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4RST-Witch-Flounder-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")
```

## Yellowtail Flounder
Otoliths from the September RV survey, from other surveys, from sentinel programs and from port sampling of commercial fisheries and from at-sea observers.

WHERE ARE THE COMMERCIAL SAMPLES FOR YEARS AFTER 1982?

```{r yellowtail, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species ==42),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV-4T"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="yelage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=42, card.type="age") ## this loads all pre-1982 commercial ages

l.4t <- which(substring(y.c$nafo.division,1,2)=="4T") ## keep only otoliths collected in NAFO 4T
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

## there is no observers otoliths for yellowtail

out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Limanda ferruginea"
out.df$english.name_nom.anglais <- "Yellowtail Flounder"
out.df$french.name_nom.français <- "Limande à queue jaune"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Yellowtail-Flounder-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

## how many were actually aged?
y <- rv$bio[which(rv$bio$species==42),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.rv.df$source <- "RV-4T"
m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## do the same for commercial samples, how many were aged
l.4tvn <- which((substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") & y.c$age %in% c(0:25)) 
comm.df2 <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.comm.df <- merge(comm.df, comm.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.comm.df$source <- "Commercial"
m.comm.df[is.na(m.comm.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

out.df <- rbind(m.rv.df, m.comm.df)
out.df$latin.name_nom.latin <- "Limanda ferruginea"
out.df$english.name_nom.anglais <- "Yellowtail Flounder"
out.df$french.name_nom.français <- "Limande à queue jaune"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Yellowtail-Flounder-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

```

## Winter Flounder
Otoliths from the September RV survey, from other surveys, from sentinel programs, from port sampling of commercial fisheries and from at-sea observers.

WHERE ARE THE COMMERCIAL SAMPLES FOR YEARS AFTER 2016?

```{r winter, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species ==43),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV-4T"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="wfage")
y.c <- gulf::read.card(year=1990, sampling="commercial", species=43, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in c(1990:1993,2012,2014:2019)){ ## loop over years
t.c <- gulf::read.card(year=yy, sampling="commercial", species=43, card.type="age")
y.c <- rbind(y.c,t.c)
}

l.4t <- which(substring(y.c$nafo.division,1,2)=="4T") ## keep only otoliths collected in NAFO 4T
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

## observers

out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Pseudopleuronectes americanus"
out.df$english.name_nom.anglais <- "Winter Flounder"
out.df$french.name_nom.français <- "Plie rouge"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Winter-Flounder-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

## how many were actually aged?
y <- rv$bio[which(rv$bio$species==43),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material==1 & y$age %in% c(0:30)),], length)
names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.rv.df$source <- "RV-4T"
m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

## do the same for commercial samples, how many were aged
l.4tvn <- which((substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") & y.c$age %in% c(0:25)) 
comm.df2 <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
m.comm.df <- merge(comm.df, comm.df2, all.x=TRUE, by=c("year_année","month_mois"))
m.comm.df$source <- "Commercial"
m.comm.df[is.na(m.comm.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0

out.df <- rbind(m.rv.df, m.comm.df)
out.df$latin.name_nom.latin <- "Pseudopleuronectes americanus"
out.df$english.name_nom.anglais <- "Winter Flounder"
out.df$french.name_nom.français <- "Plie rouge"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Winter-Flounder-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

```

## Chondrichthyan vertabrae
Vertebrae from the September RV survey.

WHERE ARE THE AGES?

```{r skates, cache=TRUE}
## RV data
y <- rv$bio[which(rv$bio$species %in% c(200:204)),]
rv.df <- aggregate(fish.number~year+month+species, data=y[which(y$age.material %in% c(1,4)),], length)
names(rv.df) <- c("year_année", "month_mois", "species", "number.vertebrae_nombre.vertèbres")
rv.df$source <- "RV-4T"
rv.df$latin.name_nom.latin <- stringr::str_to_lower(gulf::species.str(rv.df$species, "latin"))
rv.df$english.name_nom.anglais <- stringr::str_to_title(gulf::species.str(rv.df$species, "english"))
rv.df$french.name_nom.français <- gulf::species.str(rv.df$species, "french")

firstup <- function(x) {
  x <- tolower(x)
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

rv.df$latin.name_nom.latin <- firstup(rv.df$latin.name_nom.latin)

out.df <- rbind(rv.df)

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.vertebrae_nombre.vertèbres")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Chondrichthyan-otoliths.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

## how many were actually aged?
#y <- rv$bio[which(rv$bio$species %in% c(200:204)),]
#rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material %in% c(1,4)),], length)
#names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
#rv.df2 <- aggregate(fish.number~year+month, data=y[which(y$age.material %in% c(1,4) & y$age %in% c(0:30)),], length)
#names(rv.df2) <- c("year_année", "month_mois", "number.otoliths.aged_nombre.otolithes.agées")
#m.rv.df <- merge(rv.df, rv.df2, all.x=TRUE, by=c("year_année","month_mois"))
#m.rv.df$source <- "RV-4T"
#m.rv.df[is.na(m.rv.df$number.otoliths.aged_nombre.otolithes.agées),"number.otoliths.aged_nombre.otolithes.agées"] <- 0
out.df$number.otoliths.aged_nombre.otolithes.agées <- 0
vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.vertebrae_nombre.vertèbres","number.otoliths.aged_nombre.otolithes.agées")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Chondrichthyan-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE, fileEncoding = "UTF8")

```


## Atlantic Herring
Still to be integrated.
