---
title: "Gulf Fisheries Centre Ageing Structures collection"
author: "Daniel Ricard"
date: '`r paste0("Last modified timestap: ", format(Sys.time(), "%Y-%m-%d %H:%M:%S"))`'
output: html_notebook
---

The Ageing Structures collection held at the Gulf Fisheries Centre consists in Teleost saggital otoliths and Chondrichthyan vertebrae. For certain species, the physical specimens held in the collection are often accompanied with digital images. The individual details associated with each structure consists in capture date and location, length of individual and usually includes additional variables such as weight and maturity status. 

The purpose of this notebook is to facilitate the generation and update of the FGP datasets associated with the Ageing Structure collection.

```{r loaddata}
library(gulf)
data(rv)
```

## Data dictionary
```{r dictionary}
vars.bilingual <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")

data.dic <- data.frame(
  name_nom = vars.bilingual
  )
data.dic$description_fr <- c("Source de otolithes, peut soit provenir du relevé par navire de recherche (Research Vessel (RV)) ou d'échantillonnage des pêches commerciales (Commercial)", "Nom latin de l'espèce", "Nom anglais de l'espèce", "Nom français de l'espèce", "Année d'échantillonnage des otolithes", "Mois d'échantillonnage des otolithes", "Nombre d'otolithes dans la collection")

data.dic$description_en <- c("Source of otoliths, can be either from the Research Vessel (RV) survey or from commercial sampling (Commercial)", "Latin name of the species", "English name of the species", "French name of the species", "Year when otoliths were collected", "Month when otolith were collected", "Number of otoliths in collection")

data.dic.fn <- "data-dictionary.csv"
write.csv(data.dic, file=data.dic.fn, row.names = FALSE)

## data dictionary for skates
vars.bilingual <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.vertebrae_nombre.vertèbres")

data.dic <- data.frame(
  name_nom = vars.bilingual
  )
data.dic$description_fr <- c("Source des vertèbres, le relevé par navire de recherche (Research Vessel (RV))", "Nom latin de l'espèce", "Nom anglais de l'espèce", "Nom français de l'espèce", "Année d'échantillonnage des vertèbres", "Mois d'échantillonnage des vertèbres", "Nombre de vertèbres dans la collection")

data.dic$description_en <- c("Source of vertebrae, the Research Vessel (RV) survey", "Latin name of the species", "English name of the species", "French name of the species", "Year when vertebrae were collected", "Month when vertebrae were collected", "Number of vertebrae in collection")

data.dic.fn <- "data-dictionary-skates.csv"
write.csv(data.dic, file=data.dic.fn, row.names = FALSE)

```


## Atlantic Cod
Otoliths from the September RV survey, from other surveys, from sentinel programs and from commercial fisheries.
```{r cod}
## RV data
y <- rv$bio[which(rv$bio$species==10),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="codage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=10, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in c(1983:2002,2004:2018)){ ## loop over years
t.c <- gulf::read.card(year=yy, sampling="commercial", species=10, card.type="age")
y.c <- rbind(y.c,t.c)
}
l.4tvn <- which(substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") ## keep only otoliths collected in NAFO 4TVN
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Gadus morhua"
out.df$english.name_nom.anglais <- "Atlantic Cod"
out.df$french.name_nom.français <- "Morue franche"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4TVN-Atlantic-Cod-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)

```


## White Hake
Otoliths from the September RV survey, from other surveys, from sentinel programs and from commercial fisheries.
```{r hake}
## RV data
y <- rv$bio[which(rv$bio$species==12),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="whage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=12, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in c(1983:2017)){ ## loop over years
t.c <- gulf::read.card(year=yy, sampling="commercial", species=12, card.type="age")
y.c <- rbind(y.c,t.c)
}
l.4tvn <- which(substring(y.c$nafo.division,1,2)=="4T" | substring(y.c$nafo.division,1,3)=="4VN") ## keep only otoliths collected in NAFO 4TVN
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4tvn,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Urophycis tenuis"
out.df$english.name_nom.anglais <- "White Hake"
out.df$french.name_nom.français <- "Merluche blanche"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-White-Hake-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)

```

## American Plaice
Otoliths from the September RV survey, from other surveys, from sentinel programs and from commercial fisheries.
```{r plaice}
## RV data
y <- rv$bio[which(rv$bio$species==40),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="plaage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=40, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in 1983:2018){ ## loop over years
t.c <- gulf::read.card(year=yy, sampling="commercial", species=40, card.type="age")
y.c <- rbind(y.c,t.c)
}

l.4t <- which(substring(y.c$nafo.division,1,2)=="4T") ## keep only otoliths collected in NAFO 4T

comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"


out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Hippoglossoides platessoides"
out.df$english.name_nom.anglais <- "American Plaice"
out.df$french.name_nom.français <- "Plie Canadienne"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-American-Plaice-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)

```

## Witch Flounder
Otoliths from the September RV survey, from other surveys, from sentinel programs and from commercial fisheries.
```{r witch}
## RV data
y <- rv$bio[which(rv$bio$species ==41),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="witage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=41, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in 1998:2016){ ## loop over years
t.c <- gulf::read.card(year=yy, sampling="commercial", species=41, card.type="age")
y.c <- rbind(y.c,t.c)
}

l.4t <- which(substring(y.c$nafo.division,1,2) %in% c("4R","4S","4T")) ## keep only otoliths collected in NAFO 4RST
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"


out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Glyptocephalus cynoglossus"
out.df$english.name_nom.anglais <- "Witch Flounder"
out.df$french.name_nom.français <- "Plie Grise"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4RST-Witch-Flounder-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)
```

## Yellowtail Flounder
Otoliths from the September RV survey, from other surveys, from sentinel programs and from commercial fisheries.
```{r yellowtail}
## RV data
y <- rv$bio[which(rv$bio$species ==42),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="yelage")
y.c <- gulf::read.card(year=1982, sampling="commercial", species=42, card.type="age") ## this loads all pre-1982 commercial ages


l.4t <- which(substring(y.c$nafo.division,1,2)=="4T") ## keep only otoliths collected in NAFO 4T
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"

out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Limanda ferruginea"
out.df$english.name_nom.anglais <- "Yellowtail Flounder"
out.df$french.name_nom.français <- "Limande à queue jaune"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Yellowtail-Flounder-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)
```

## Winter Flounder
Otoliths from the September RV survey, from other surveys, from sentinel programs and from commercial fisheries.
```{r winter}
## RV data
y <- rv$bio[which(rv$bio$species ==43),]
rv.df <- aggregate(fish.number~year+month, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
rv.df$source <- "RV"

## commercial data
## what years are available for commercial ages?
## com.files <- list.files(path="//ENT.dfo-mpo.ca/dfo-mpo/GROUP/GLF/Regional_Shares/Science/Hd2/commercial/age", pattern="wfage")
y.c <- gulf::read.card(year=1990, sampling="commercial", species=43, card.type="age") ## this loads all pre-1982 commercial ages
for(yy in c(1990,1991,1992,1993,2012,2014,2015,2016)){ ## loop over years
t.c <- gulf::read.card(year=yy, sampling="commercial", species=43, card.type="age")
y.c <- rbind(y.c,t.c)
}

l.4t <- which(substring(y.c$nafo.division,1,2)=="4T") ## keep only otoliths collected in NAFO 4T
comm.df <- aggregate(otolith.number~year+month, data=y.c[l.4t,], length)
names(comm.df) <- c("year_année", "month_mois", "number.otoliths_nombre.otolithes")
comm.df$source <- "Commercial"


out.df <- rbind(rv.df, comm.df)
out.df$latin.name_nom.latin <- "Pseudopleuronectes americanus"
out.df$english.name_nom.anglais <- "Winter Flounder"
out.df$french.name_nom.français <- "Plie Rouge"

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.otoliths_nombre.otolithes")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Winter-Flounder-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)

```

## Chondrichthyan vertabrae
Vertebrae from the September RV survey.

```{r skates}
## RV data
y <- rv$bio[which(rv$bio$species %in% c(200:204)),]
rv.df <- aggregate(fish.number~year+month+species, data=y[which(y$age.material==1),], length)
names(rv.df) <- c("year_année", "month_mois", "species", "number.vertebrae_nombre.vertèbres")
rv.df$source <- "RV"
rv.df$latin.name_nom.latin <- stringr::str_to_lower(gulf::species.str(rv.df$species, "latin"))
rv.df$english.name_nom.anglais <- stringr::str_to_title(gulf::species.str(rv.df$species, "english"))
rv.df$french.name_nom.français <- gulf::species.str(rv.df$species, "french")

firstup <- function(x) {
  x <- tolower(x)
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

rv.df$latin.name_nom.latin <- firstup(rv.df$latin.name_nom.latin)

out.df <- rbind(rv.df)

vars <- c("source", "latin.name_nom.latin", "english.name_nom.anglais", "french.name_nom.français", "year_année","month_mois","number.vertebrae_nombre.vertèbres")
o1 <- order(out.df$source, out.df$year_année, out.df$month_mois)

out.fn <- "./NAFO-4T-Chondrichthyan-ages.csv"
write.csv(out.df[o1,vars], file=out.fn, row.names = FALSE)

```



## Atlantic Herring

